### 2020-06-22 Multiobjective-Optimization-in-Image-Approximation/paper.tex
The paper title was selected.

Multiobjective Optimization in Image Approximation