### 2020-06-23 Multiobjective-Optimization-in-Image-Approximation/paper.tex
Descriptive information about the article has been added.

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{Multiobjective Optimization in Image Approximation \\ {}
\thanks{The publication has been prepared with the support of Velbazhd Software LLC and Bulgarian Ministry of Education and Science according to the research project No. {D01–205/23.11.2018}.}}

\author{\IEEEauthorblockN{Plamen Petrov}
\IEEEauthorblockA{\textit{Institute of Information and Communication Technologies} \\
\textit{Bulgarian Academy of Sciences} \\
1113 Sofia, Bulgaria \\
p.petrov@iit.bas.bg}
\and
\IEEEauthorblockN{Georgi Kostadinov}
\IEEEauthorblockA{\textit{Institute of Information and Communication Technologies} \\
\textit{Bulgarian Academy of Sciences} \\
1113 Sofia, Bulgaria \\
g.kostadinov@iit.bas.bg}
\and
\IEEEauthorblockN{Petar Zhivkov}
\IEEEauthorblockA{\textit{Institute of Information and Communication Technologies} \\
\textit{Bulgarian Academy of Sciences} \\
1113 Sofia, Bulgaria \\
City, Country \\
pzhivkov@iit.bas.bg}
\and
\IEEEauthorblockN{Veneta Velichkova}
\IEEEauthorblockA{\textit{Institute of Information and Communication Technologies} \\
\textit{Bulgarian Academy of Sciences} \\
1113 Sofia, Bulgaria \\
vvelichkova@iit.bas.bg}
\and
\IEEEauthorblockN{Todor Balabanov}
\IEEEauthorblockA{\textit{Institute of Information and Communication Technologies} \\
\textit{Bulgarian Academy of Sciences} \\
1113 Sofia, Bulgaria \\
0000-0003-3139-069X}
}

%
% Plamen Dimitrov Petrov - p.petrov@iit.bas.bg
% Georgi Borisov Kostadinov - g.kostadinov@iit.bas.bg
% Petar Rumenov Zhivkov - pzhivkov@iit.bas.bg
% Veneta Ivanova Velichkova - vvelichkova@iit.bas.bg
% Todor Dimitrov Balabanov - todorb@iinf.bas.bg
%
% Bulgarian Academy of Sciences
% Institute of Information and Communication Technologies
% acad. Georgi Bonchev Str., block 2, office 514
% 1113 Sofia, Bulgaria
% http://iict.bas.bg/
%

\maketitle

\begin{abstract}
\end{abstract}

\begin{IEEEkeywords}
image approximation, genetic algorithms, colors reduction, image vectorization
\end{IEEEkeywords}

\section{Introduction}

\section{Conclusion}

\section*{Acknowledgment}

This research is funded by Velbazhd Software LLC and it is partially supported by the Bulgarian Ministry of Education and Science (contract D01–205/23.11.2018) under the National Scientific Program ``Information and Communication Technologies for a Single Digital Market in Science, Education and Security (ICTinSES)'', approved by DCM \# 577/17.08.2018.

\begin{thebibliography}{00}

\bibitem{}

\end{thebibliography}

\end{document}